---
title: "P8105 Homework 1"
author: "sarah8105"
date: "2020-09-14"
output: github_document
---


This is my first homework assignment for P8105. This assignment demonstrates the use of git and R Markdown documents.


## Problem 1

The following code chunk creates a dataframe comprised of:

* A random sample of size **10** from a standard Normal distribution
* A logical vector indicating whether elements of the sample are greater than 0
* A character vector of length **10**
* A factor vector of length **10**, with factors "low", "middle", or "high"

The code chunk then attempts to take the mean of each variable.

```{r problem1_a}
library(tidyverse)

problem1_df = tibble(
  vec_numeric = rnorm(10),
  vec_logical = ifelse(vec_numeric > 0, TRUE, FALSE),
  vec_character = c("This", "is", "a", "character", "vector", "for", "the", "first", "homework", "problem"),
  vec_factor = factor(c("low", "middle", "high", "low", "low", "middle", "high", "low", "middle", "middle"))
)

mean(problem1_df$vec_numeric)

mean(problem1_df$vec_logical)

mean(problem1_df$vec_character)

mean(problem1_df$vec_factor)
```

The mean of the random numeric variable is `r round(mean(problem1_df$vec_numeric), 2)` and the mean of the logical variable is `r mean(problem1_df$vec_logical)`. Means are unable to be calculated from the character or factor variables since these variables cannot be implicitly coerced into numeric variables.

The following second code chunk attempts to convert the logical, character, and factor vectors to numeric and multiplies the result by the random sample.

```{r problem1_b, results = 'hide'}
as.numeric(problem1_df$vec_logical)

vec_log_to_num = as.numeric(problem1_df$vec_logical) * problem1_df$vec_numeric

as.numeric(problem1_df$vec_character)

vec_char_to_num = as.numeric(problem1_df$vec_character) * problem1_df$vec_numeric

as.numeric(problem1_df$vec_factor)

vec_fact_to_num = as.numeric(problem1_df$vec_factor) * problem1_df$vec_numeric

```

The factor and logical vectors are successfully converted to numeric vectors (with _TRUE = 1_ and _FALSE = 0_ and with _low = 2_, _middle = 3_ and _high = 1_ respectively). 

R implicitly coerces logical variables to numeric when taking the mean. Using the as.numeric function can explicitly convert a factor variable to its underlying numeric code. The character vector cannot be implicitly or explicitly converted to numeric, so a mean cannot be calculated.

## Problem 2

Problem 2 downloads the `penguins` dataset from the `palmerpenguins` package.